---
title: "eda"
author: "Anand Rajan"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown
d that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
weather_df %>%
  group_by(name, month)

weather_df %>% 
  group_by(name,month) %>% 
  summarize(n_obs=n())

weather_df %>% 
  pull(name) %>% 
  table()
```

it's nice that summarize produces a data frame

'table ' doesnt, its a bad function

```{r}
weather_df %>% 
  janitor::tabyl(month,name)
```

## More General summaries
Let's count, take means, look at SDs..

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax)) 


weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

Formatting Tables
knitr kable allows you to format table nicely

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```


##Grouped Mutate

group specific centering NOT residual

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
```


Lag: is what happened in previous observations affecting current observations

for example air pollution, its all relative

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))


weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n=1),
    tmax_diff = tmax - lagged_tmax
  ) %>% 
  summarize(diff_sd=sd(tmax_diff, na.rm=TRUE))
  
```


## Limitations


```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax = cor(tmin,tmax,use="complete"))

weather_df %>%
  group_by(name) %>%
  mutate(
    temp_change = tmax - lag(tmax),
    tmax_diff=tmax-temp_change
         ) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

