---
title: "visualization1"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

## R Markdown

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots

You can also embed plots, for example:

```{r}
ggplot(weather_df, aes(x=tmin, y = tmax)) + geom_point()
```

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax))+ geom_point()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x=tmin, y=tmax))

weather_plot + geom_point()
```
##Advanced Scatterplot

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name)) + 
  geom_point() +
  geom_smooth(se= FALSE)

## what about 'aes' placement..?

``{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax)) + 
  geom_point(aes(color=name)) + 
  geom_smooth()

```


## faceit_grid (rows ~ columns)
alpha can be both mapped to a specific variable(within the ggplot function) or it can be done globally (withing geom_point)
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color=name)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth(se=FALSE, size = 2) +
  facet_grid(. ~ name)
```


combine some elements and try a new plot

```{r}
weather_df %>%  
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point(aes(size=prcp),alpha = 0.5) +
  geom_smooth()+
  facet_grid(. ~ name)
  

```

## some small notes
how many geoms have to exist?

You can have whatever geoms you want. 

```{r}

weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_smooth(se=FALSE)
```

you can use  a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax)) +
  geom_density2d()+
  geom_point()
```

## Univariate Plots
Histograms 

```{r}
weather_df %>% 
  ggplot(aes(x=tmin)) +
  geom_histogram()
```

add color...

```{r}

weather_df %>% 
  ggplot(aes(x=tmin)) +
  geom_histogram() + 
  facet_grid(. ~ name)
```

If I want to compare distributions across different levels, use different geometry!

```{r}

weather_df %>% 
  ggplot(aes(x=tmin, fill=name)) +
  geom_density(alpha=0.3, adjust = 0.5)
```
Density is a smoothed out histogram around the edges

box plots

```{r}
weather_df %>% 
  ggplot(aes(x=name, y=tmin)) +
  geom_boxplot()
```


```{r}
weather_df %>% 
  ggplot(aes(x=name, y=tmin, fill=name)) +
  geom_violin(alpha=0.5) +
  stat_summary(fun="median")
```

Ridge plots 

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=name)) +
  geom_density_ridges()

```

